---
title: "MS5130 A3 Clothing_Product_Analysis"
author: "Manav"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
toc: true
---

## Product analysis using Dataset 1

```{r}

# Loading all the required libraries

library(ggplot2) # For visualization
library(dplyr) # For data manipulation
library(tidyr) # For tidying data
library(stringr) # For string manipulation
library(readr) # For reading CSV files
library(tidytext)
library(plotly)
library(wordcloud)
library(quanteda)
```

```{r}

# Reading the CSV File
product_catalog <- read_csv("C:/Users/manav/OneDrive/Desktop/Work/R_Applied_Analytics/Assignment 3/myntra_products_catalog.csv")

nrow(product_catalog)

# Pre-processing data
product_catalog <- product_catalog %>%
  # Renaming columns named weirdly
  rename(PriceINR = `Price (INR)`) %>%
  # Removing rows with blank entries or NAs in any column
  ####filter(apply(., 1, function(row) all(row != "" & !is.na(row)))) %>%
  drop_na() %>%
  # Trimming leading and trailing whitespaces in all character columns
  # Adapted from stackoverflow: https://stackoverflow.com/questions/20760547/removing-whitespace-from-a-whole-data-frame-in-r
  mutate(across(where(is.character), trimws))

head(product_catalog)


```

### Expensive and budget-friendly brands

```{r}

# Calculating the average price for each brand
brand_prices <- product_catalog %>%
  group_by(ProductBrand) %>%
  # Adapted from stackoverflow : https://stackoverflow.com/questions/25759891/dplyr-summarise-each-with-na-rm
  summarise(AveragePrice = mean(PriceINR, na.rm = TRUE),NumProducts = n()) %>%
  filter(NumProducts > 10) %>%
  ungroup() %>%
  arrange(desc(AveragePrice))

# Identifying the top 3 most expensive brands and cheap brands
top_expensive_brands <- head(brand_prices, 3)

top_cheap_brands <- tail(brand_prices, 3)

# Displaying the results
print("Top 3 Most Expensive Brands: ") 
print(top_expensive_brands)

print("Top 3 Budget Friendly Brands: ")
print(top_cheap_brands)


```

### Gender wise product distribution

```{r}

# Getting product counts grouped by gender
gender_counts <- product_catalog %>%
  group_by(Gender) %>%
  summarise(Count = n()) %>%
  ungroup()

# Creating a bar graph for the number of products by gender
ggplot(gender_counts, aes(x = Gender, y = Count, fill = Gender, label = Count)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  theme_minimal() +
  labs(title = "Gender wise product distribution",
       x = "Gender",
       y = "Number of Products") +
  scale_fill_brewer(palette = "Set1") +
  geom_text(position = position_stack(vjust = 0.5))


```

### Color wise product distribution

```{r}

# Getting product counts for each color
color_counts <- product_catalog %>%
  group_by(PrimaryColor) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))
nrow(color_counts)

# Identifying top 5 colors
top_colors <- head(color_counts, 5)


# Putting the rest of the colors in "Others" category
others_count <- sum(tail(color_counts, n = nrow(color_counts) - 5)$Count)
others_row <- data.frame(PrimaryColor = "Others", Count = others_count)

# Combining the top colors with "Others"
final_data <- rbind(top_colors, others_row)

final_data

# setting color styles manually
# Adapted from open ai : 'Prompt - Color codes for the below colors'
named_colors <- c(
  "Black" = "#808080",  # Lighter grey instead of black
  "Blue" = "#ADD8E6",   # Light Blue
  "Red" = "#FF9999",    # Light Red
  "Green" = "#90EE90",  # Light Green
  "White" = "#F8F8FF",  # Off-White, Ghost White
  "Others" = "#D8BFD8"  # Light Purple (Thistle)
)

# Changing the 'final_data' values as percentage of whole count
final_data$Percentage <- (final_data$Count / sum(final_data$Count)) * 100

ggplot(final_data, aes(x = "", y = Count, fill = PrimaryColor, label = sprintf("%1.1f%%", Percentage))) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Color Based Product Distribution") +
  scale_fill_manual(values = named_colors) +
  geom_text(position = position_stack(vjust = 0.5))

# blue, black and red are the most popular color categories
```

## Sentiment and Review analysis using Dataset 2 and 3

```{r}

# Reading the other 2 datasets (Detailed Product dataset and Product Review dataset)

Product_Detailed_2 <- read_csv("C:/Users/manav/OneDrive/Desktop/Work/R_Applied_Analytics/Assignment 3/Product_Detailed_2.csv")
Product_Reviews_3 <- read_csv("C:/Users/manav/OneDrive/Desktop/Work/R_Applied_Analytics/Assignment 3/Product_Reviews_3.csv")

# Manually creating the mapping based on domain knowledge so that the Categories from both the dataset can be matched to each other

Product_Detailed_2 <- Product_Detailed_2 %>%
  mutate(Category = case_when(
    Category == "Inner Wear & Sleep Wear" ~ "Intimate",
    Category == "Lingerie & Sleep Wear" ~ "Intimate",
    Category == "Western" ~ "Dresses",
    Category == "Bottom Wear" ~ "Bottoms",
    Category == "Topwear" ~ "Tops",
    Category == "Sports Wear" ~ "Jackets",
    Category == "Indian Wear" ~ "Trend",
    TRUE ~ as.character(Category)  # Default case to keep original Category values
  ))
```

### Sentiment scores and average ratings

```{r}

# Average rating using Detailed product dataset (Dataset 2)

avg_rating_by_category <- Product_Detailed_2 %>%
  group_by(Category) %>%
  summarise(AverageRating = mean(Ratings, na.rm = TRUE)) %>%
  ungroup()

# Sentiment scores using Product Review dataset (Dataset 3)

sentiment_analysis <- Product_Reviews_3 %>%
  unnest_tokens(word, 'Review Text') %>%
  inner_join(get_sentiments("bing"), by = "word", relationship = "many-to-many") %>%
  group_by(index) %>%
  summarise(sentiment_score = sum(case_when(
    sentiment == "positive" ~ 1,
    sentiment == "negative" ~ -1,
    TRUE ~ 0L
  )), .groups = 'drop')


# plotting sentiment score histogram

ggplot(sentiment_analysis, aes(x = sentiment_score)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Sentiment Scores", x = "Sentiment Score", y = "Count")


# Joining the sentiments back to the review dataset
reviewdataset_with_sentiments <- Product_Reviews_3 %>%
  left_join(sentiment_analysis, by = "index")

# Grouping by Category for analysis
avg_sentiments_by_category <- reviewdataset_with_sentiments %>%
  group_by(Category) %>%
  summarise(average_sentiment = mean(sentiment_score, na.rm = TRUE)) %>%
  ungroup()  

# Visualizing sentiment scores across categories
avg_sentiments_by_category <- na.omit(avg_sentiments_by_category)

p <- plot_ly(data = avg_sentiments_by_category, x = ~Category, y = ~average_sentiment, type = 'bar', marker = list(color = ~average_sentiment, colorscale = 'Viridis'), showlegend = FALSE) %>%
  layout(title = 'Distribution of Average Sentiment Scores by Category',
         xaxis = list(title = 'Category'),
         yaxis = list(title = 'Average Sentiment Score'))

p

```

### Insights using Sentiment score and ratings

```{r}

# Joining the datasets on the 'Category' column and displaying the results
category_insights <- inner_join(avg_rating_by_category, avg_sentiments_by_category, by = "Category")

final_display <- category_insights %>%
  select(Category, AverageRating, average_sentiment)

print(final_display)

# The bar chart
p <- plot_ly(data = final_display, x = ~Category) %>%
  add_bars(y = ~AverageRating, name = 'Average Rating', marker = list(color = '#ADD8E6')) %>%
  add_bars(y = ~average_sentiment, name = 'Average Sentiment', marker = list(color = '#FF9999')) %>%
  layout(yaxis = list(title = 'Score'),
         barmode = 'group',
         title = 'Comparison of Average Ratings and Sentiments by Category')

p
```

The bar chart shows comparison of average ratings and average sentiments across various categories such as "Bottoms," "Dresses," "Intimate," and so on.

**Insights from the chart:**

Consistency: The average ratings and average sentiments are consistent across categories. This suggests that customers’ overall satisfaction levels are similar across different product types. Moreover the balance indicates that sentiment scores derived from reviews correlate well with the numerical ratings customers give. In other words, how customers speak about the product aligns with how they rate it.

Lack of Extremes and positiveness: There are no categories with extreme sentiment or rating scores. The positive ratings and sentiment values also indicate a moderately positive level of customer feeling towards the products in each categories.

### Wordcloud of Category with high sentiment

```{r}

# Filtering the dataset for the category "Tops"
dresses_reviews <- Product_Reviews_3 %>%
  filter(Category == "Tops") %>%
  select('Review Text')

# Creating a text corpus from the vector source of review text
text_corpus <- corpus(dresses_reviews$'Review Text')

# Preprocessing to remove punctuations, numbers, spaces and common english words like 'the', 'at' since they are not relevant for analysis
cleaned_corpus <- text_corpus %>%
  tokens() %>%
  tokens_tolower() %>%
  tokens_remove(pattern = "[[:punct:]]") %>%
  tokens_remove(pattern = "\\d+") %>%
  tokens_remove(pattern = stopwords("en")) %>%
  tokens_remove(pattern = "[[:space:]]+")

# Generate the word cloud
set.seed(1234)
wordcloud(words = cleaned_corpus, scale = c(3, 0.5), max.words = 100, random.order = FALSE, colors = brewer.pal(8, "Dark2"))

```

The word cloud for "tops" highlights key words that stand out in customer reviews. "Love," "like," and "fit" are prominent, indicating these are common sentiments that customers frequently mention. Words like "color," "size," and "fabric" suggest these are important factors for customers when evaluating tops. Overall, this word cloud reflects positive customer experiences.
