---
title: "MS5130 A3 Clothing_Product_Analysis"
author: "Manav"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
toc: true
---

## Product analysis using Dataset 1

```{r}

# Loading all the required libraries

library(ggplot2) # For visualization
library(dplyr) # For data manipulation
library(tidyr) # For tidying data
library(stringr) # For string manipulation
library(readr) # For reading CSV files
library(tidytext) 
library(plotly)
library(wordcloud)
library(quanteda)
library(leaflet)
library(tidygeocoder)

```

```{r}

# Reading the CSV File
product_catalog <- read_csv("myntra_products_catalog.csv")

nrow(product_catalog)

# Pre-processing data
product_catalog <- product_catalog %>%
  # Renaming columns named weirdly
  rename(PriceINR = `Price (INR)`) %>%
  # Removing rows with blank entries or NAs in any column
  ####filter(apply(., 1, function(row) all(row != "" & !is.na(row)))) %>%
  drop_na() %>%
  # Trimming leading and trailing whitespaces in all character columns
  # Adapted from stackoverflow: https://stackoverflow.com/questions/20760547/removing-whitespace-from-a-whole-data-frame-in-r
  mutate(across(where(is.character), trimws))

head(product_catalog)


```

### Expensive and budget-friendly brands

```{r}

# Calculating the average price for each brand
brand_prices <- product_catalog %>%
  group_by(ProductBrand) %>%
  # Adapted from stackoverflow : https://stackoverflow.com/questions/25759891/dplyr-summarise-each-with-na-rm
  summarise(AveragePrice = mean(PriceINR, na.rm = TRUE),NumProducts = n()) %>%
  filter(NumProducts > 10) %>%
  ungroup() %>%
  arrange(desc(AveragePrice))

# Identifying the top 3 most expensive brands and cheap brands
top_expensive_brands <- head(brand_prices, 3)

top_cheap_brands <- tail(brand_prices, 3)

# Displaying the results
print("Top 3 Most Expensive Brands: ") 
print(top_expensive_brands)

print("Top 3 Budget Friendly Brands: ")
print(top_cheap_brands)


```

### Gender wise product distribution

```{r}

# Getting product counts grouped by gender
gender_counts <- product_catalog %>%
  group_by(Gender) %>%
  summarise(Count = n()) %>%
  ungroup()

# Creating a bar graph for the number of products by gender
ggplot(gender_counts, aes(x = Gender, y = Count, fill = Gender, label = Count)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  theme_minimal() +
  labs(title = "Gender wise product distribution",
       x = "Gender",
       y = "Number of Products") +
  scale_fill_brewer(palette = "Set1") +
  geom_text(position = position_stack(vjust = 0.5))


```

### Color wise product distribution

```{r}

# Getting product counts for each color
color_counts <- product_catalog %>%
  group_by(PrimaryColor) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))
nrow(color_counts)

# Identifying top 5 colors
top_colors <- head(color_counts, 5)


# Putting the rest of the colors in "Others" category
others_count <- sum(tail(color_counts, n = nrow(color_counts) - 5)$Count)
others_row <- data.frame(PrimaryColor = "Others", Count = others_count)

# Combining the top colors with "Others"
final_data <- rbind(top_colors, others_row)

final_data

# setting color styles manually
# Adapted from open ai : 'Prompt - Color codes for the below colors'
named_colors <- c(
  "Black" = "#808080",  # Lighter grey instead of black
  "Blue" = "#ADD8E6",   # Light Blue
  "Red" = "#FF9999",    # Light Red
  "Green" = "#90EE90",  # Light Green
  "White" = "#F8F8FF",  # Off-White, Ghost White
  "Others" = "#D8BFD8"  # Light Purple (Thistle)
)

# Changing the 'final_data' values as percentage of whole count
final_data$Percentage <- (final_data$Count / sum(final_data$Count)) * 100

ggplot(final_data, aes(x = "", y = Count, fill = PrimaryColor, label = sprintf("%1.1f%%", Percentage))) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Color Based Product Distribution") +
  scale_fill_manual(values = named_colors) +
  geom_text(position = position_stack(vjust = 0.5))

# blue, black and red are the most popular color categories
```

## Sentiment and Review analysis using Dataset 2 and 3

```{r}

# Reading the other 2 datasets (Detailed Product dataset and Product Review dataset)

Product_Detailed_2 <- read_csv("Product_Detailed_2.csv")
Product_Reviews_3 <- read_csv("Product_Reviews_3.csv")

# Manually creating the mapping based on domain knowledge so that the Categories from both the dataset can be matched to each other

Product_Detailed_2 <- Product_Detailed_2 %>%
  mutate(Category = case_when(
    Category == "Inner Wear & Sleep Wear" ~ "Intimate",
    Category == "Lingerie & Sleep Wear" ~ "Intimate",
    Category == "Western" ~ "Dresses",
    Category == "Bottom Wear" ~ "Bottoms",
    Category == "Topwear" ~ "Tops",
    Category == "Sports Wear" ~ "Jackets",
    Category == "Indian Wear" ~ "Trend",
    TRUE ~ as.character(Category)  # Default case to keep original Category values
  ))
```

### Sentiment scores and average ratings

```{r}

# Average rating using Detailed product dataset (Dataset 2)

avg_rating_by_category <- Product_Detailed_2 %>%
  group_by(Category) %>%
  summarise(AverageRating = mean(Ratings, na.rm = TRUE)) %>%
  ungroup()

# Sentiment scores using Product Review dataset (Dataset 3)

sentiment_analysis <- Product_Reviews_3 %>%
  unnest_tokens(word, 'Review Text') %>%
  inner_join(get_sentiments("bing"), by = "word", relationship = "many-to-many") %>%
  group_by(index) %>%
  summarise(sentiment_score = sum(case_when(
    sentiment == "positive" ~ 1,
    sentiment == "negative" ~ -1,
    TRUE ~ 0L
  )), .groups = 'drop')


# plotting sentiment score histogram

ggplot(sentiment_analysis, aes(x = sentiment_score)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Sentiment Scores", x = "Sentiment Score", y = "Count")


# Joining the sentiments back to the review dataset
reviewdataset_with_sentiments <- Product_Reviews_3 %>%
  left_join(sentiment_analysis, by = "index")

# Grouping by Category for analysis
avg_sentiments_by_category <- reviewdataset_with_sentiments %>%
  group_by(Category) %>%
  summarise(average_sentiment = mean(sentiment_score, na.rm = TRUE)) %>%
  ungroup()  

# Visualizing sentiment scores across categories
avg_sentiments_by_category <- na.omit(avg_sentiments_by_category)

p <- plot_ly(data = avg_sentiments_by_category, x = ~Category, y = ~average_sentiment, type = 'bar', marker = list(color = ~average_sentiment, colorscale = 'Viridis'), showlegend = FALSE) %>%
  layout(title = 'Distribution of Average Sentiment Scores by Category',
         xaxis = list(title = 'Category'),
         yaxis = list(title = 'Average Sentiment Score'))

p

```

The histogram shows a normal distribution for the sentiment scores, most scores cluster around the middle, suggesting that most sentiments expressed are neutral. There are fewer very positive or very negative sentiments, as seen by the lower bars at the ends.

This bar chart displays average sentiment scores for different clothing categories. Tops, Dresses, and Jackets seem to have higher average sentiments, indicating more positive feedback, while the 'Trend' category has a noticeably lower average sentiment score.

### Insights using Sentiment score and ratings

```{r}

# Joining the datasets on the 'Category' column and displaying the results
category_insights <- inner_join(avg_rating_by_category, avg_sentiments_by_category, by = "Category")

final_display <- category_insights %>%
  select(Category, AverageRating, average_sentiment)

print(final_display)

# The bar chart
p <- plot_ly(data = final_display, x = ~Category) %>%
  add_bars(y = ~AverageRating, name = 'Average Rating', marker = list(color = '#ADD8E6')) %>%
  add_bars(y = ~average_sentiment, name = 'Average Sentiment', marker = list(color = '#FF9999')) %>%
  layout(yaxis = list(title = 'Score'),
         barmode = 'group',
         title = 'Comparison of Average Ratings and Sentiments by Category')

p
```

The bar chart shows comparison of average ratings and average sentiments across various categories such as "Bottoms," "Dresses," "Intimate," and so on.

**Insights from the chart:**

Consistency: The average ratings and average sentiments are consistent across categories. This suggests that customers’ overall satisfaction levels are similar across different product types. Moreover the balance indicates that sentiment scores derived from reviews correlate well with the numerical ratings customers give. In other words, how customers speak about the product aligns with how they rate it.

Lack of Extremes and positiveness: There are no categories with extreme sentiment or rating scores. The positive ratings and sentiment values also indicate a moderately positive level of customer feeling towards the products in each categories.

### Wordcloud of Category with high sentiment

```{r}

# Filtering the dataset for the category "Tops"
dresses_reviews <- Product_Reviews_3 %>%
  filter(Category == "Tops") %>%
  select('Review Text')

# Creating a text corpus from the vector source of review text
text_corpus <- corpus(dresses_reviews$'Review Text')

# Preprocessing to remove punctuations, numbers, spaces and common english words like 'the', 'at' since they are not relevant for analysis
cleaned_corpus <- text_corpus %>%
  tokens() %>%
  tokens_tolower() %>%
  tokens_remove(pattern = "[[:punct:]]") %>%
  tokens_remove(pattern = "\\d+") %>%
  tokens_remove(pattern = stopwords("en")) %>%
  tokens_remove(pattern = "[[:space:]]+")
# Adapted from: https://search.r-project.org/CRAN/refmans/quanteda/html/tokens_select.html

# Generate the word cloud
set.seed(1234)
wordcloud(words = cleaned_corpus, scale = c(3, 0.5), max.words = 100, random.order = FALSE, colors = brewer.pal(8, "Dark2"))

```

The word cloud for "tops" highlights key words that stand out in customer reviews. "Love," "like," and "fit" are prominent, indicating these are common sentiments that customers frequently mention. Words like "color," "size," and "fabric" suggest these are important factors for customers when evaluating tops. Overall, this word cloud reflects positive customer experiences.

## Analysis of Brand Average Ratings in Relation to Pricing and Product Count

```{r}

# Aggregating brand data using Dataset 1 - product_catalog

agg_product_data <- product_catalog %>%
  group_by(ProductBrand) %>%
  filter(n() > 50) %>%
  summarise(
    AvgPrice1 = mean(PriceINR, na.rm = TRUE),
    ProductCount = n()
  ) %>%
  ungroup()

# Aggregating brand data using Dataset 2 - Product_Detailed_2

agg_detailed_data <- Product_Detailed_2 %>%
  group_by(BrandName) %>%
  filter(n() > 50) %>%
  summarise(
    AvgPrice2 = mean(`OriginalPrice (in Rs)`, na.rm = TRUE),
    AvgRating = mean(Ratings, na.rm = TRUE)
  ) %>%
  ungroup()

combined_data <- merge(agg_product_data, agg_detailed_data, 
                       by.x = "ProductBrand", by.y = "BrandName")

head(combined_data)

# Building and summarizing a glm model

glm_model <- glm(AvgRating ~ AvgPrice1 + ProductCount, family = gaussian(), data = combined_data)

summary(glm_model)

```

A Generalized Linear Model (GLM) explores the relationships between the average price and the product count on the average rating for brands with more than 50 products.

The GLM analysis, focusing on the average price and the product count as predictors of average rating, reveals that neither average price nor product count significantly affects the average ratings in a statistically meaningful way.

```{r}

# Adding predicted ratings
combined_data$predicted_ratings <- predict(glm_model, type = "response")

# Ploting the ratings data
ggplot(combined_data, aes(x = AvgRating, y = predicted_ratings)) +
  geom_point(aes(color = "Actual vs. Predicted"), alpha = 0.6) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs. Predicted Ratings",
       x = "Actual Ratings",
       y = "Predicted Ratings") +
  theme_minimal() +
  scale_color_manual(values = c("Actual vs. Predicted" = "blue"))


```

While the statistical model did not reveal significant predictors for average ratings, the visualization of actual versus predicted ratings is presented here to show the model's power.

We can see that most points are clustered around the line, but there is a spread, indicating some variance between the predicted and actual ratings.

## Average price changes over time

We are using Top 5 categories from the 'brand_prices' variable generated in 'Expensive and budget-friendly brands' section that uses **product dataset (dataset 1)** that has products from **2015** and also generating top 5 brands from the detailed dataset **(dataset 2)** that has products from **2020** for brands that have more than 10 products and then comparing the prices over time.

```{r}

# Aggregating brand data using Dataset 1 - product_catalog
top5_expensive_brands_2015 <- head(brand_prices, 5) %>%
  mutate(Year = "2015") %>%
  select(Brand = ProductBrand, AvgPrice=AveragePrice, Year)

# Aggregating brand data using Dataset 2 - Product_Detailed_2

top5_expensive_brands_2020 <- Product_Detailed_2 %>%
  group_by(BrandName) %>%
  filter(n() > 10) %>%
  summarise(
    AvgPrice = mean(`OriginalPrice (in Rs)`, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(Year = "2020") %>%
  arrange(desc(AvgPrice)) %>%
  head(5) %>%
  select(Brand = BrandName, AvgPrice, Year)


# Binding the two while keeping the year distinction
top_brands_combined <- bind_rows(top5_expensive_brands_2015, top5_expensive_brands_2020)

top_brands_combined


# Arranging the dataframe by AvgPrice in descending order

top_brands_combined <- top_brands_combined %>%
  mutate(Brand = factor(Brand, levels = unique(Brand[order(-AvgPrice)])))

# Creating the plot and also displaying top 5 expensive brands from 2015 and 2020

df_2015 <- top_brands_combined %>% filter(Year == "2015")
df_2020 <- top_brands_combined %>% filter(Year == "2020")

print("Top 5 Expensive brands 2015:") 
df_2015

print("Top 5 Expensive brands 2020:")
df_2020

# Creating the bar chart for 2015
p1 <- plot_ly(data = df_2015, x = ~Brand, y = ~AvgPrice, type = 'bar', name = '2015',
              marker = list(color = 'coral')) %>%
  layout(yaxis = list(title = 'Average Price 2015 (INR)'))

# Creating the bar chart for 2020
p2 <- plot_ly(data = df_2020, x = ~Brand, y = ~AvgPrice, type = 'bar', name = '2020',
              marker = list(color = 'seagreen')) %>%
  layout(yaxis = list(title = 'Average Price 2020 (INR)'),
         xaxis = list(title = 'Brand'))

# Combining the two plots into one with two subplots
# Note: shareX = TRUE means that all subplots will share the same x-axis.
subplot(p1, p2, nrows = 2, shareX = TRUE) %>%
  layout(title = 'Average Prices by Brand for 2015 and 2020')
# Adapted from: https://plotly.com/r/subplots/

```

This output shows a side-by-side comparison of average prices for various brands in two different years, 2015 and 2020. The orange bars represent the average prices in 2015, while the green bars represent the prices in 2020. The chart shows that, in general, brands are charging more for their products in 2020 than they did in 2015, prices have gone up in five years. The top brands have also changed like 'Teakwood', 'White Fire', however brands like 'Seiko' continue to dominate the expensive segment.

## Brand location on world map using leaflet

```{r}

# Loading the dataset that contains brands along with their founding country
location_of_brands <- read.csv("Location_of_brands.csv")

head(location_of_brands)

# Finding lat and long using OSM geocoding
geocoded_data <- location_of_brands %>%
  geocode(country = location, method = 'osm') %>%
 # Adapted from: https://geocoder.readthedocs.io/providers/OpenStreetMap.html
  distinct(location, .keep_all = TRUE) %>%
  select(location,lat,long)

print(geocoded_data)

# Joining the lat and long with the brands with location dataset
brand_location_geocoded <- inner_join(location_of_brands, geocoded_data, by = "location", relationship = "many-to-many") %>%
  select (Brands,location,lat,long)

head(brand_location_geocoded)

# Aggregating the data from detailed dataset 2 at brand level
aggregated_brands <- Product_Detailed_2 %>%
  group_by(BrandName) %>%
  summarise(
    AvgPrice = mean(`OriginalPrice (in Rs)`, na.rm = TRUE),
    ProductCount = n()
  ) %>%
  ungroup() %>%
  select(Brands = BrandName, AvgPrice, ProductCount)


# Using the Brand aggregated dataset and brand geocoded dataset to create a list of top 3 brands for each country and product count for those brands

top_brands_by_country <- aggregated_brands %>%
  inner_join(brand_location_geocoded, by = "Brands") %>%
  group_by(location) %>%
  arrange(desc(AvgPrice)) %>%
  mutate(Rank = row_number()) %>%
  filter(Rank <= 3) %>%
  summarise(
    TopBrands = paste(Brands, "(", round(AvgPrice, 2), "INR - ", ProductCount, "products)", collapse = "<br>")
  ) %>%
  ungroup()

# Joining back with 'geocoded_data' to get latitude and longitude
country_info <- geocoded_data %>%
  distinct(location, .keep_all = TRUE) %>%
  inner_join(top_brands_by_country, by = "location")

print(country_info)

# Creating the world map using leaflet
leaflet(country_info) %>%
  addProviderTiles(providers$CartoDB.Voyager) %>%
  # Adapted from: https://www.kaggle.com/code/devzohaib/interactive-maps-with-leaflet-in-r
  addCircleMarkers(
    lng = ~long, lat = ~lat, radius = 8, color = "coral", fillOpacity = 0.8,
    popup = ~paste0("<strong>", location, "</strong><br><strong>Top 3 Brands:</strong><br>", TopBrands)
  ) %>%
  setView(lng = 20, lat = 0, zoom = 2)

```
